---
title: "Exploring Iris Pitch"
author: "Scott D. Weitzenhoffer"
date: "February 6, 2018"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Iris Data

This famous (Fisher's or Anderson's) iris data set gives the measurements in
centimeters of the variables sepal length and width and petal length and width,
respectively, for 50 flowers from each of 3 species of iris. The species are
Iris setosa, versicolor, and virginica.

###Source

_Fisher, R. A. (1936) The use of multiple measurements in taxonomic problems. 
Annals of Eugenics, 7, Part II, 179–188._

_The data were collected by Anderson, Edgar (1935). The irises of the Gaspe 
Peninsula, Bulletin of the American Iris Society, 59, 2–5._


## Concept

You've extensively explored this Iris data set in R.

Wouldn't you like to share your love of this data set with other people who
don't have R installed on their system?

## Plotly and Shiny

Using Shiny and Plotly, you can deliver to your audience a beautiful 3D plot of
iris data, Your users can pick which of Sepal Length, Sepal Width, Petal Length,
and Petal Width to plot, and produce plots like this:

```{r message=FALSE}
library(plotly)
p <- plot_ly(x=iris$Sepal.Length, 
             y=iris$Sepal.Width, 
             z=iris$Petal.Length,
             color=iris$Species,
             marker = list(size = 2)) %>%
    layout(title="Iris data: 3D Chart",
           scene=list(
               xaxis=list(title="Sepal Length"),
               yaxis=list(title="Sepal Width"),
               zaxis=list(title="Petal Length"),
               autosize=FALSE,
               width=250,
               height=250
           )
    )
p
```

## Try it today!

Head over to [Exploring Iris](https://sweitzen.shinyapps.io/ExploringIris/),
and try it for yourself!

Exploring Iris works on Chrome and Internet Explorer, but not on Firefox.